---
title: "Day four - 7/13"
output: 
  html_document:
  keep_md: yes
author: "Marcus W. Beck, beck.marcus@epa.gov"
---

### Editing metadata in R

Use `library(EML)`

```{r}
library(EML)
# library(arcticdatautils) # eml_creator

doc <- read_eml('strix-pacific-northwest.xml')
doc
```

Edit the creator field:
```{r eval = F}
doc@dataset@creator
doc@dataset@creator <- c(eml_creator('Turbo', 'Todd'), email = 'turbotodd@rad.com')
```
Edit the abstract, write file.
```{r, eval = F}
# Change abstract to the better one we wrote
doc@dataset@abstract <- as(set_TextType("better-abstract.md"), "abstract")

# Save it back to the filesystem
write_eml(doc, "strix-pacific-northwest.xml")
```

### Publishing data (on DataONE)

Published as a data package which is one or more datasets and a metadata file.  

An example:
```{r,eval = T}
library(dataone)
library(datapack)
library(uuid)

# create an empty package
dp <- new("DataPackage")
dp
```

We then add a metadata file:
```{r}
emlFile <- "strix-pacific-northwest.xml"
metadataObj <- new("DataObject", format="eml://ecoinformatics.org/eml-2.1.1", filename=paste(getwd(), emlFile, sep="/"),
                   suggestedFilename=basename(emlFile))
dp <- addData(dp, metadataObj)
```

We then add data:
```{r eval = F}
sourceData <- "sample.csv"
sourceObj <- new("DataObject", format="text/csv", filename=paste(getwd(), sourceData, sep="/"), suggestedFilename=basename(sourceData))
dp <- addData(dp, sourceObj, metadataObj)
```

Then maybe add some script for processing the data:
```{r eval = F}
progFile <- "filterSpecies.R"
progObj <- new("DataObject", format="application/R", filename=paste(getwd(), progFile, sep="/"), mediaType="text/x-rsrc",
                suggestedFilename=basename(progFile))
dp <- addData(dp, progObj, metadataObj)
```

Then the processed data file is added:
```{r eval = F}
outputData <- "filteredSpecies.csv"
outputObj <- new("DataObject", format="text/csv", filename=paste(getwd(), outputData, sep="/"), suggestedFilename=basename(outputData))
dp <- addData(dp, outputObj, metadataObj)

myAccessRules <- data.frame(subject="http://orcid.org/0000-0003-0077-4738", permission="changePermission") 

dp
```

Finally, we upload the package to the Testing server for the KNB.
```{r, eval=FALSE}
d1c <- D1Client("STAGING2", "urn:node:mnTestKNB")
packageId <- uploadDataPackage(d1c, dp, public=TRUE, accessRules=myAccessRules, quiet=FALSE)
```

This particular package contains the R script `filterSpecies.R`, the input file `sample.csv` that was read by
the script and the output file `filteredSpecies.csv` that was created by the R script, which was run at a
previous time.  

You can now search for and view the package at https://dev.nceas.ucsb.edu

**Make DataONE profile to doc and upload data!!**

### Open data in R

Go here to check out list of open data sources that can be dled from within R:

[https://github.com/ropensci/opendata](https://github.com/ropensci/opendata)

Examples:

`rplos`: R client for the PLoS Journals API

- Package: https://github.com/ropensci/rplos
- Website: http://api.plos.org/

```{r}
library(rplos)
searchplos(q='everything:"gulf of mexico"', fl='title', fq='doc_type:full', limit=10)
```

`rnaturalearth`: an R package to hold and facilitate interaction with natural earth map data **useful for geopolitical boundaries!!**

- Website: www.naturalearthdata.com/
- Package: https://github.com/ropenscilabs/rnaturalearth

`rnoaa`: R interface to many NOAA data APIs 

Access data like:

- Air temps
- Sea ice extent
- Buoy data
- Tons more!

```{r}
library(rnoaa)

# Go here: http://www.ndbc.noaa.gov/
# Find a station ID, like http://www.ndbc.noaa.gov/station_page.php?station=42039
bd <- buoy(dataset = "cwind",  buoyid = 42039, datatype = "cc")
plot(bd$data$wind_spd)
```

### Web Scraping 

[Look here!](https://nceas.github.io/oss-lessons/data-liberation/intro-webscraping.html)

Low-level packages: `Rcurl`, `httr` for basic requests (get, give)

More user-friendly packages: `rvest`, `rselenium` (clicking buttons!!)

Request to server, response from client - language

HTML - blocks/tags in a hierarcy

CSS - formatting HTML when it's rendered.  

Selector-gadget...

```{r}
library("rvest")

URL <- "http://santabarbaraca.com/explore-and-discover-santa-barbara/neighborhoods-towns/santa-barbara/the-funk-zone/"

# Read the webpage into R
webpage <- read_html(URL)

# Parse the webpage for bars
# use selector gadget to identify nodes!!!!!!!!!!!!!
bars <- html_nodes(webpage, ".neighborhoods-towns-business .neighborhoods-towns-business-title")

# Extract the name of the bar
bar_names <- html_text(bars)

```

Scrape some GIS files:

```{r}
url <- 'http://sero.nmfs.noaa.gov/maps_gis_data/fisheries/gom/GOM_index.html'

shps <- url %>% 
  read_html %>% 
  html_nodes(".dataBox a") %>% 
  html_attr('href') %>% 
  grep('\\.zip$', ., value = T) %>% 
  paste0('http://sero.nmfs.noaa.gov', .)

shps
```
