---
title: "Day two - 7/11"
output: 
  html_document:
    keep_md: yes
author: "Marcus W. Beck, beck.marcus@epa.gov"
---
```{r, echo = F}
library(knitr)
opts_chunk$set(warning = F)
```

### Version control: git basics

Distributed - you have the entire history of a git project on your local machine, applies to everyone working separately on a the same project. 

See the git manual:
```{r}
system('git', intern = TRUE)
```

See your global configurations:
```{r}
system('git config --global --list', intern = TRUE)
```

Save your user credentials:
```{r eval = F}
# forever
system('git config --global credential.helper cache', intern = TRUE)

# for an hour
system("git config --global credential.helper 'cache --timeout=3600'", intern = TRUE)

```

Show hidden files using `ls -a`, show files recursively `ls -R`. Look into cygwin or commander for a good windows shell.  

`git clone` copy a remote repo to your local machine (or server)

`vim favorite_desserts.csv` creates a file and enters the interactive text editor.  Enter text as needed.  To exit, press esc, `:w` to save, `:q` to quit.  
 
 Or enter text from command line. 
 ```
 echo 'line 1, line 1' > favorite_desserts.csv
 ```
 To append to a file, use two carets `>>`.
 
 `git diff` shows what's changed since last commit. 
 
 `git checkout -- filename` will revert a file back to its version in a previous commit.
 
 Putting a git repository on Aurora, login to Aurora with new Git Bash shell:
 `ssh beck@aurora.nceas.ucsb.edu` (to exit server execute `exit`)
 
 Then run: `git clone https://github.com/fawda123/myrepo.git` in whatever directory on Aurora. 
 
 To see a file tree in Git Bash `cmd //c tree`
